---
title: "2020 Elections Data"
author: "Vivian Zhang"
date: "4/10/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(devtools)
library(fivethirtyeight)
library(gov.1005.data)
library(ggthemes)

# I uploaded all of these packages. Initially, I didn't have ggthemes and fivethirtyeight, but I realized that I needed those for the graphs.

```

```{r madlibs, echo=FALSE}
q1 <- endorsements_2020 %>%
  arrange(date) %>%
  slice(1) %>%
  pull(date)

# I rearranged the data in endorsements, sliced to take the first row only, the pulled only from the column named date. I learned that slice gives me the number of rows I ask for and that pull gives me the column I want.

q2 <- endorsements_2020 %>%
  arrange(desc(date)) %>%
  slice(1) %>%
  pull(date)

# Same idea as the first question. This time, I arranged the data in descending order based on date so that the most recent date comes first. Then I sliced and pulled so I could get the date value of the first row that comes up.

q3 <- endorsements_2020 %>%
  filter(position == "mayor" & 
           endorsee == "Pete Buttigieg") %>%
  pull(endorser)

# I filtered based on two criteria: mayor and pete. Initially, I wasn't sure I could place both of these inside the same filter function. Katherine initially had it written under two filter functions. However, this is the simpler way to write the code.

q4 <- endorsements_2020 %>%
  filter(state == "NH") %>%
  summarize(mean(points))

# I filtered for New Hampshire results only, then used the summarize functioning, taking the mean result only.

q5 <- endorsements_2020 %>%
  filter(endorsee == "Amy Klobuchar" | 
           endorsee == "Elizabeth Warren") %>%
  summarize(sum(points))

# The vertical bar means or. Everything else follows the same ideas as the previous questions. We use summarize and filter functions.

qa <- endorsements_2020 %>%
  pull("position") %>%
  class()

# I made a mistake initially because I didn't realize that the class function wasn't supposed to have anything inside as input.

qb <- endorsements_2020 %>%
  pull("endorser") %>%
  class()

# This is the same ideas as the position, simply pulling endorser instead of position.

qc <- endorsements_2020 %>%
  pull("points") %>%
  class()

# This is again the same ideas as above, but we simply pull points instead of endorser.

```

1. The earliest date for an endorsement in the dataset is `r q1`.

2. The most recent date for an endorsement is `r q2`.

3. `r q3` is the mayor that has endorsed Pete Buttigieg.

4. Endorsers from New Hampshire have an average of `r q4` endorsement points.

5. Between them, the two female candidates who remain in the race (Amy Klobuchar and Elizabeth Warren) have a total of `r q5` endorsement points.

6.

a. The variable “position” is a `r qa` class of variable.

b. The variable “endorser” is a `r qb` class of variable.

c. The variable “points” is a `r qc` class of variable.

```{r graph, echo=FALSE}

avg_data <- congress %>%
  filter(party %in% c("D", "R")) %>%
  group_by(party, year) %>%
  summarise(avg_age = mean(age))

# Above, we are filtering the congress data and grouping it by party and year. We then summarize it, taking the mean. This will give us the average age of people in a given party in a given year.

ggplot(avg_data, aes(x = year, y = avg_age, color = party)) + 
  geom_line() + 
  labs(title = "Average Age of Members of Congress", 
       subtitle = "At start of term, 1947-2013",
       caption = "Source: fivethirtyeight package") + 
  scale_color_manual(values = c("blue", "red")) + theme_fivethirtyeight() + 
  annotate("text", x = 1966, y = 57, 
           label = "Republicans", color = "red", size = 4.5) + 
  annotate("text", x = 1996.5, y = 58, 
           label = "Democrats", color = "blue", size = 4.5) +
  scale_x_continuous(limits = c(1947, 2014),
                     breaks = c(1950, 1960, 1970, 1980, 1990, 2000, 2010), labels = c("1950", "'60", "'70", "'80", "'90", "2000", "'10" )) + 
  scale_y_continuous(limits = c(40, 62), breaks = 
                       c(40, 45, 50, 55, 60), labels = c("40", "45", "50", "55", "60 yrs")) +
  theme(legend.position = "none")

# Here, we are using ggplot and setting the x axis to year, y axis to average age. The color changes based on the party. This was helpful because I learned to use the labs function. This is what we use to add titles, subtitles, and captions. I added the scale_color_manual function, which allows us to assign specific colors to each of the parties. I added the theme_fivethirtyeight() so that the formatting of the graph is more similar to the example. Then, I annotated the actual graph by setting different coordinates and labeling Democrats vs Republicans. Writing scale_x_continuous and scale_y_continuous was a bit tedious because we have to list out all the breaks and labels that we want to have. Finally, I got rid of the legend by typing theme(legend.position = "none).

```